\documentclass[11pt]{article}
% To produce a letter size output. Otherwise will be A4 size.
\usepackage[letterpaper]{geometry}
% For enumerated lists using letters: a. b. etc.
\usepackage{enumitem}
\topmargin -.5in
\textheight 9in
\oddsidemargin -.25in
\evensidemargin -.25in
\textwidth 7in
\begin{document}
% Edit the following putting your first and last names and your lab section.
\author{Tiffany Pham\\
Lab CSE-015-08L F 8:30-10:20pm}
% Edit the following replacing X with the HW number.
\title{CSE 015: Discrete Mathematics\\
Spring 2024\\
Homework \#02\\
Solution}
% Put today's date in the following.
\date{February 22, 2024}
\maketitle
% ========== Begin questions here
\begin{enumerate}
    \item \textbf{Question 1:}
    \begin{enumerate}[label=(\alph*)]
        \item If an animal is a rabbit, then every rabbit hops.
        \item Every animal is a rabbit and hops.
        \item There exists an animal such that if it is a rabbit, then it hops.
        \item There exists an animal that is a rabbit and hops.
    \end{enumerate}
    \item \textbf{Question 2:}
    \begin{enumerate}[label=(\alph*)]
        \item True (let $x=\sqrt{2}$)
        \item False ($\sqrt{-1}$ is not a real number)
        \item True (the left hand side is always at least 2)
        \item False (not true for $x=1$ or $x=0$)
    \end{enumerate}
    \item \textbf{Question 3:}
    \begin{enumerate}[label=(\alph*)]
        \item Let $P(x)$ be "$x$ has a cell phone." Then we have $\forall P(x)$ the first way, or $\forall x(C(x)\rightarrow P(x))$ the second way.
        \item Let $F(x)$ be "$x$ has seen a foreign movie." Then we have $\exists xF(x)$ the first way, or $\exists(C(x)\land F(x))$ the second way.
        \item Let $S(x)$ be "$x$ can swim." Then we have $\exists x\lnot S(x)$ the first way, or $\exists x(C(x)\land\lnot S(x))$ the second way.
        \item Let $Q(x)$ be "$x$ can solve quadratic equations." Then we have $\forall xQ(X)$ the first way, or $\forall x(C(x)\rightarrow Q(x))$ the second way.
    \end{enumerate}
    \item \textbf{Question 4:}
    \begin{enumerate}[label=(\alph*)]
        \item Some student in your class has taken some computer science course.
        \item Every student in your class has taken at least one computer science course.
        \item Every computer science course has been taken by at least one student in your class.
    \end{enumerate}
    \item \textbf{Question 5:}
    \begin{enumerate}[label=(\alph*)]
        \item $\exists x\exists yQ(x,y)$
        \item $\lnot\exists x\exists yQ(x,y)$ or $\forall x\forall y\lnot Q(x,y)$.
    \end{enumerate}
    \item \textbf{Question 6:}
    \begin{enumerate}[label=(\alph*)]
        \item $\forall x\exists yF(x,y)$
        \item $\lnot\exists x\forall yF(x,y)$
        \item $\forall y\exists xF(x,y)$
    \end{enumerate}
    \item \textbf{Question 7:}
    \begin{enumerate}[label=(\alph*)]
        \item $\forall x\forall y(((x<0)\land(y<0))\rightarrow(xy>0))$
        \item $\forall x\forall y(((x>0)\land(y>0))\rightarrow(((x+y)/2>0))$
        \item $\forall x\forall y(|x+y|\leq|x|+|y|)$
    \end{enumerate}
    \item \textbf{Question 8:}
    \begin{enumerate}[label=(\alph*)]
        \item False
        \item True
        \item False
    \end{enumerate}
\end{enumerate}
\end{document}
